---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Transform Bio-Oracle NetCDF files into GeoTIFF

<!-- badges: start -->

<!-- badges: end -->

Bio-Oracle layers are natively developed in NetCDF format. This short tutorial shows how to turn into GeoTIFF in R and Python.

As an example, we will download the [surface Ocean Temperature layer from Bio-Oracle](https://bio-oracle.org/downloads-to-email-v3.php), predicted for the decade of 2090 to 2100 using the Shared Socioeconomic Pathway scenario ssp585. Here below you can find a direct link to download.

[**Download Bio-Oracle Ocean Temperature Layer**](https://erddap.bio-oracle.org/erddap/griddap/thetao_ssp585_2020_2100_depthsurf.nc?thetao_mean[(2090-01-01):1:(2090-01-01T00:00:00Z)][(-90.0):1:(90.0)][(-180.0):1:(180.0)])

Save the layer in a folder named `data` and name it as `bo.nc`.

# NetCDF to GeoTIFF in R

There are several ways of turning a NetCDF file into GeoTIFF. Here we show how to do it with two of the most common packages for dealing with raster data.

## Using terra

```{r terra}
# Install required package
# You can install them using: install.packages("terra")
library(terra)

# Path to the Bio-Oracle layer file
nc_path <- "./data/bo.nc"

# Read layer
bo <- rast(nc_path)

# Inspect file
bo

# Plot
plot(bo)

# Write as GeoTIFF
writeRaster(bo, "./data/bo_terra.tif")
```

## Using stars

```{r stars}
# Install required package
# You can install them using: install.packages("stars")
library(stars)

# Path to the Bio-Oracle layer file
nc_path <- "./data/bo.nc"

# Read layer
bo <- read_stars(nc_path)

# Inspect file
bo

# Plot
plot(bo)

# Write as GeoTIFF
write_stars(bo, "./data/bo_stars.tif")
```


# NetCDF to GeoTIFF in Python

```{python xarray}
# Install required packages
# You can install them using: pip install xarray rioxarray matplotlib
import xarray as xr
import rioxarray
import matplotlib.pyplot as plt

# Path to the Bio-Oracle layer file
nc_path = "./data/bo.nc"

# Read layer
bo = xr.open_dataset(nc_path)

# Inspect file
print(bo)

# Extract the variable data that you want to convert to GeoTIFF
variable_data = bo["thetao_mean"]

# Plot
variable_data.plot()
plt.show()

# Write as GeoTIFF using rioxarray
variable_data.rio.to_raster("./data/bo_xarray.tif")
```